Computing Entropies with Nested Sampling

### 2017-12-08 talk/ms-abridged.tex
Removed some more stuff

% New slide
\begin{frame}
\frametitle{Interpretation of conditional entropies}

How uncertain would the question ``what's the value of $\theta$,
precisely?'' be if the question ``what's the value of $d$, precisely''
were to be resolved?

% New slide
\begin{frame}
\frametitle{Connections}

Entropy of the joint prior:
\begin{align}
H(\theta, d) &= H(\theta) + H(d | \theta) \\
                          &= H(d) + H(\theta | d).
\end{align}

Mutual information:
\begin{align}
I(\theta; d) &= H(\theta) - H(\theta | d).
\end{align}
This quantifies dependence --- or more fundamentally,
relevance, or the potential for learning.

{\tiny There are many other ways of expressing $I$.}

@@

% New slide
\begin{frame}
\frametitle{Hard integrals}
Hard integrals with nuisance parameters
$\boldsymbol{\eta}$, interesting parameter(s)
$\boldsymbol{\phi}$
\vspace{0.5em}

\begin{center}
\includegraphics[width=0.9\textwidth]{hard-integrals.png}
\end{center}

@@

or
\begin{align}
Z &= \int \pi(\theta) L(\theta) \, d\theta
\end{align}
where $\pi=$ prior, $L=$ likelihood.
It's just an expectation. Why is it hard?
\end{frame}

% New slide
\begin{frame}
\frametitle{Simple Monte Carlo fails}

\begin{align}
Z &= \int \pi(\theta) L(\theta) \, d\theta \\
  &\approx \frac{1}{N} \sum_{i=1}^N L(\theta_i)
\end{align}
with $\theta_i \sim \pi$ will probably miss the
tiny regions where $L$ is high.

Equivalently, $\pi$ implies a {\em very heavy-tailed}
distribution of $L$-values, and simple Monte Carlo fails.

@@

% New slide
\begin{frame}
\frametitle{Nested Sampling}
The idea is to generate a sequence of points with increasing
likelihoods, such that we can estimate their $X$ values. Since
we know their $L$-values, we can then do the integral 
numerically.

\begin{align}
Z = \int_0^1 L(X) dX
\end{align}

% New slide
\begin{frame}
\frametitle{Nested Sampling algorithm}

\begin{enumerate}
  \item<2-> Generate $N$ points from $\pi$
  \item<3-> Find the worst one (lowest likelihood $L^*$), save it
  \item<4-> Estimate its $X$ using Beta$(N,1)$ distribution\footnote{From order statistics.}
  \item<5-> Record worst particle and its $X$-value, then discard it.
  \item<6-> Replace that point with a new one from $\pi$ but with
        likelihood above $L^*$.
  \item<7-> Repeat steps 2--5 indefinitely.
\end{enumerate}
\end{frame}

@@

% New slide
\begin{frame}
\frametitle{In practice...}

When a period is short relative to observations, you can get a multimodal
posterior pdf\footnote{e.g., see Larry Bretthorst's work connecting
the posterior pdf to the periodogram.},
and `learn a lot' by ruling out most of the space, but still
having many peaks.\vspace{0.5em}

I did not investigate this aspect of the problem --- I assumed long periods.

@@

{\color{blue}\url{http://www.mdpi.com/1099-4300/19/8/422}
\url{https://github.com/eggplantbren/InfoNest}}
\vspace{0.5cm}

>>>>

@@

It's even worse with nuisance parameters.

@@

or
\begin{align}
Z &= \int \pi(\theta) L(\theta) \, d\theta
\end{align}
where $\pi=$ prior, $L=$ likelihood.
This is hard because $\pi \implies$ a heavy-tailed
distribution of $L$-values.

@@

@@

@@

{\color{blue}\bf \url{http://www.mdpi.com/1099-4300/19/8/422}
\url{https://github.com/eggplantbren/InfoNest}}
\vspace{0.5cm}
